---
title: "Classwork October 2"
author: "Marguerite CAMERON"
date: "2 octobre 2017"
output: html_document
---

---
title: "Week 2 notes"
output:
  html_document: default
  pdf_document: default
---

```{r, echo=F}
# THIS IS THE VERSION I WROTE INITIALLY
plot_hist <- function(d, title) {
  p <- ggplot2::ggplot(d, ggplot2::aes(x=logRT)) +
    ggplot2::geom_histogram(fill=fill_colour, colour="black") +
    ggplot2::ylim(c(0,0.15)) +
    ggplot2::ggtitle(title)
  print(p)
}

# THIS IS THE MORE ADVANCED VERSION
plot_hist_fancy <- function(d, title,
                      variable="logRT",
                      fill_colour="lightgrey",
                      median_line=T,
                      line_at=NULL) {
  p <- ggplot2::ggplot(d,
                       ggplot2::aes_string(x=variable,
                                    y="(..count..)/sum(..count..)")) +
    ggplot2::geom_histogram(fill=fill_colour, colour="black") +
    ggplot2::ylim(c(0,0.15)) +
    ggplot2::ggtitle(title)
  if (median_line) {
    p <- p + ggplot2::geom_vline(ggplot2::aes(xintercept=median(logRT)),
                                 lwd=3)
  }
  if (!is.null(line_at)) {
    p <- p + ggplot2::geom_vline(xintercept=line_at,
                                 colour="red",
                                 lwd=2)    
  }
  print(p)
}

randomize_rts <- function(d) {
  result <- dplyr::select(d, Condition, logRT)
  result$logRT <- sample(result$logRT, nrow(d))
  return(result)
}

fraction_above <- function(x, criterion) {
  n_above <- sum(x > criterion)
  n <- length(x)
  return(n_above/n)
}

fraction_below <- function(x, criterion) {
  n_below <- sum(x <= criterion)
  n <- length(x)
  return(n_below/n)
}

fraction_between <- function(x, minimum, maximum) {
  n_between <- sum(x > minimum & x <= maximum)
  n <- length(x)
  return(n_between/n)
}

difference_in_medians <- function(d) {
  d_gram <- dplyr::filter(d, Condition == "Grammatical")
  d_ungram <- dplyr::filter(d, Condition == "Ungrammatical")
  return(median(d_gram$logRT) - median(d_ungram$logRT))
}

difference_in_fraction <- function(d, minimum=6.5, maximum=7) {
  d_gram <- dplyr::filter(d, Condition == "Grammatical")
  d_ungram <- dplyr::filter(d, Condition == "Ungrammatical")
  return(fraction_between(d_gram$logRT, minimum, maximum) -
           fraction_between(d_ungram$logRT, minimum, maximum))
}

sum_of_group <- function(d, group="Grammatical") {
  d_group <- dplyr::filter(d, Condition == group)
  return(sum(d_group$logRT))
}

```

## Basic plots and descriptive statistics

The reaction time data is first transformed using log.

```{r}
spr <- missingverb::spr_english
spr$logRT <- log(spr$RT)
```

Histogram of log reaction times in grammatical and ungrammatical  conditions.

```{r, out.width="400px"}
spr_gram <- dplyr::filter(spr, Condition == "Grammatical")
spr_ungram <- dplyr::filter(spr, Condition == "Ungrammatical")
plot_hist(spr_gram, "Grammatical")
plot_hist(spr_ungram, "Ungrammatical")
```

The number of data points in the grammatical condition is `r nrow(spr_gram)`; the number of data points in the ungrammatical condition is `r nrow(spr_ungram)`.

Half of the data points in the grammatical condition are below **`r median(spr_gram$logRT)`**; half of the data points in the ungrammatical condition are below **`r median(spr_ungram$logRT)`**. The difference in the medians is **`r difference_in_medians(spr)`**.

The fraction of the data that is between 6.5 and 7 in the grammatical condition is **`r mean(spr_gram$logRT >= 6.5 & spr_gram$logRT < 7)`**; the fraction of the data that is between 6.5 and 7 in the ungrammatical condition is **`r mean(spr_ungram$logRT >= 6.5 & spr_ungram$logRT < 7)`**. The difference is **`r difference_in_fraction(spr, 6.5, 7)`**

The overall total in the grammatical condition is **`r sum(spr_gram$logRT)`**; the overall total in the ungrammatical condition is **`r sum(spr_ungram$logRT)`**.


## Hypothetical version of the data where the two conditions aren't different

Data plotted all together:

```{r}
plot_hist(spr, "Pooled")
```

Fake data:

```{r, out.width="300px"}
spr_fake <- randomize_rts(spr)
spr_fake_gram <- dplyr::filter(spr_fake, Condition == "Grammatical")
spr_fake_ungram <- dplyr::filter(spr_fake, Condition == "Ungrammatical")
plot_hist(spr_fake_gram, "Grammatical (fake)")
plot_hist(spr_fake_ungram, "Ungrammatical (fake)")
```

The number of data points in the fake grammatical condition is **`r nrow(spr_fake_gram)`**; the number of data points in the fake ungrammatical condition is **`r nrow(spr_fake_ungram)`**.

Half of the data points in the grammatical condition are below **`r median(spr_fake_gram$logRT)`**; half of the data points in the ungrammatical condition are below **`r median(spr_fake_ungram$logRT)`**. The difference in the medians is **`r difference_in_medians(spr_fake)`**.

The fraction of the data that is between 6.5 and 7 in the grammatical condition is **`r mean(spr_fake_gram$logRT >= 6.5 & spr_fake_gram$logRT < 7)`**; the fraction of the data that is below 7 in the ungrammatical condition is **`r mean(spr_fake_ungram$logRT >= 6.5 & spr_fake_ungram$logRT < 7)`**. The difference is **`r difference_in_fraction(spr_fake, 6.5, 7)`**

The overall total in the grammatical condition is **`r sum(spr_fake_gram$logRT)`**; the overall total in the ungrammatical condition is **`r sum(spr_fake_ungram$logRT)`**.


## 1000 hypothetical versions of the data where the two conditions aren't different

Create a thousand fake samples.

```{r, cache=T}
N_SAMPLES <- 1000
stat1_difference_in_medians <- rep(0, N_SAMPLES)
stat2_difference_in_fraction <- rep(0, N_SAMPLES)
stat3_total_grammatical <- rep(0, N_SAMPLES)
stat4_total_ungrammatical <- rep(0, N_SAMPLES)
for (i in 1:N_SAMPLES) {
  spr_fake <- randomize_rts(spr)
  spr_fake_gram <- dplyr::filter(spr_fake, Condition == "Grammatical")
  spr_fake_ungram <- dplyr::filter(spr_fake, Condition == "Ungrammatical")
  stat1_difference_in_medians[i] <- difference_in_medians(spr_fake)
  stat2_difference_in_fraction[i] <- difference_in_fraction(spr_fake, 6.5, 7)
  stat3_total_grammatical[i] <- sum(spr_fake_gram$logRT)
  stat4_total_ungrammatical[i] <- sum(spr_fake_ungram$logRT)
}
statistics <- tibble::tibble(Medians=stat1_difference_in_medians,
                             Fractions=stat2_difference_in_fraction,
                             TotalGram=stat3_total_grammatical,
                             TotalUngram=stat4_total_ungrammatical)
```

Plot the distribution of the four statistics.

```{r}
plot_hist_fancy(statistics, "Difference in medians", "Medians", median_line=F,
          fill_colour="#FFD86D",
          line_at=difference_in_medians(spr))
plot_hist_fancy(statistics, "Difference in % between 6.5 and 7", "Fractions", 
          median_line=F, fill_colour="#FFD86D",
          line_at=difference_in_fraction(spr, 6.5, 7))
plot_hist_fancy(statistics, "Total log(RT) - grammatical", "TotalGram", 
          median_line=F, fill_colour="#FFD86D",
          line_at=sum(spr_gram$logRT))
plot_hist_fancy(statistics, "Total log(RT) - ungrammatical", "TotalUngram", 
          median_line=F, fill_colour="#FFD86D",
          line_at=sum(spr_ungram$logRT))
```

**The observed difference in medians is bigger than `r fraction_below(statistics$Medians, difference_in_medians(spr))*100.`% of the samples.**

**The observed difference in the fraction of observations between 6.5 and 7 is bigger than `r fraction_below(statistics$Fractions, difference_in_fraction(spr, 6.5, 7))*100.`% of the samples.**

**The observed overall total logRT in the grammatical cases is bigger than `r fraction_below(statistics$TotalGram, sum(spr_gram$logRT))*100.`% of the samples.**

**The observed overall total logRT in the ungrammatical cases is smaller than `r fraction_above(statistics$TotalUngram, sum(spr_ungram$logRT))*100.`% of the samples.**