---
title: "Missing Verb"
author: "Marguerite CAMERON"
date: "2 octobre 2017"
output: html_document
---

## Basic plots and descriptive statistiques 

```{r}

```


```{r}
 spr <- missingverb::spr_english
spr$logRT <- log(spr$RT)
hist(spr$logRT)
```

```{r}
spr_gram <- dplyr::filter(spr, Condition == "Grammatical")
spr_ungram <- dplyr::filter(spr, Condition == "Ungrammatical")
hist(spr_gram$logRT)
```

```{r}
plot_hist <- function(d, title) {
    p <- ggplot2(d, ggplot2::aes(x=logRT, 
                                 y=(..count..)/sum(..count..)))
    p <- p + ggplot2::geom_histogram(fill="lightgrey", 
                                    color="black")
    p <- p + ggplot2::ggtitle(title)
    print(p)
}
spr_gram <- dplyr::filter(spr, Condition == "Grammatical")
spr_ungram <- dplyr::filter(spr, Condition == "Ungrammatical")
plot_hist(spr_gram, "Title")
plot_hist(spr_gram, "Ungrammatical Condition")
```
## Hypothetical version of the data where the two conditions aren't different

```{r}
plot_hist(spr, "Pooled")
```
Fake data:

```{r}
spr_fake <- spr
print(spr_fake)
```


The number of data points in the real grammatical condition is
'nrow(spr_gram)'. The number of data points in the real ungrammatical condition is 'nrow(spr_ungram)'.

Half of the data points in the grammatical condition are below 


## 200 Hypothetical versions of the data where the two conditions aren't different

```{r, cache=T}
N_SAMPLES <- 200
statistics <- rep(0, N_SAMPLES)
for (i in 1:N_SAMPLES) {
  spr_fake <- dplyr::filter(spr_fake, Condition == "Grammatical")
  spr_fake_gram <- dplyr::filter(spr_fake, Condition == "Ungrammatical")
  statistics[i] <- median(spr_fake_gram$logRT) - 
                    median(spr_fake_ungram$logRT)
}
statistics_d <- tibble::tibble(MedianDifference=statistics
)
print(statistics_d)
plot_hist_fancy(statistics_d, 
                "Sampling distribution of difference in medians",
                "MedianDifference")
```

